### 2013-08-22 doc/dissertation/document/backmatter/parallel_metrics.tex
removing redundant fanm verification data.

How effectively the presented algorithms parallelize MCSA will be
determined by several performance measures. For this work, these
measures will come directly from Keyes' primer on the scalability of
domain decomposed methods and specifically his discussion on how to
measure their parallel performance \cite{keyes_how_1999}. All of
these metrics are derived from two measured quantities: the number of
iterations required for convergence and the total wall time required
for convergence. Each of these quantities are then measured in varying
parameter studies by changing the number of processes in a simulation
and by varying the local or global problem size. Using the notation of
Keyes, we will define \textit{parallel efficiency} as $\eta(N,P)$
where the efficiency is a function of the global problem size, $N$,
and the number of processes in the calculation, $P$. There are several
types of efficiencies that will be useful in our analysis of
scalability with different definitions depending on the type of
parametric scaling study performed. For all efficiencies, a value of
100\% is considered perfect while a value of 0\% signals the worst
possible performance. Often, as was observed in this work, it is the
case that efficiencies above 100\% are computed in certain scaling
studies. For the data presented in the following sections, an
explanation of each observed case of super-unitary efficiencies will
be provided.

>>>>

How effectively the algorithms presented in this work parallelize MCSA
will be determined by several performance measures. For this work,
these measures will come directly from Keyes' primer on the
scalability of domain decomposed methods and specifically his
discussion on how to measure their parallel performance
\cite{keyes_how_1999}. All of these metrics are derived from two
measured quantities: the number of iterations required for convergence
and the total wall time required for convergence. Each of these
quantities are then measured in varying parameter studies by changing
the number of processes in a simulation and by varying the local or
global problem size. Using the notation of Keyes, we will define
\textit{parallel efficiency} as $\eta(N,P)$ where the efficiency is a
function of the global problem size, $N$, and the number of processes
in the calculation, $P$. There are several types of efficiencies that
will be useful in our analysis of scalability with different
definitions depending on the type of parametric scaling study
performed. For all efficiencies, a value of 100\% is considered
perfect while a value of 0\% signals the worst possible
performance. Often, as was observed in this work, it is the case that
efficiencies above 100\% are computed in certain scaling studies. For
the data presented in the following sections, an explanation of each
observed case of super-unitary efficiencies will be provided.

### 2013-08-22 doc/dissertation/document/chapters/nonlinear_problem/nonlinear_problem.tex
removing redundant fanm verification data.

%%---------------------------------------------------------------------------%%
\section{The FANM Method\ }
\label{sec:fanm}
In production physics codes based on nonlinear equations sets,
Newton-Krylov methods are the primary means of generating a fully
consistent solution scheme
\cite{evans_development_2006,evans_enhanced_2007,gaston_parallel_2009,godoy_parallel_2012}. It
is common that for large scale simulations these problems are memory
limited due to the subspaces generated by robust Krylov methods which
may often build hundreds of vectors during a Newton step. Often, a
matrix-free approach is chosen to relax memory requirements over
directly generating the Jacobian matrix and facilitate the
implementation. However, as we observed in previous sections, these
matrix-free methods suffer from poorly scaled problems and the first
order error introduced by the Jacobian approximation. In addition, it
was observed that the savings induced by the matrix-free approach is
eventually amortized over a number of nonlinear iterations where it
becomes more efficient computationally to instead form the Jacobian.

@@

For each of the benchmark problems, we will present the results of
computations using FANM and directly compare them to results using a
Newton-Krylov method in order to verify the correctness of FANM and
its applicability to fluid problems. For each case, we will compare
the global minimum, maximum, and average values for fluid velocity and
pressure for all benchmarks and additionally temperature for the
thermal convection cavity problem. If the values for these quantities
computed using FANM match those computed using a production
Newton-Krylov method, the implementation will be deemed correct and
the method deemed valid for the purposes of this work. In addition, we
will report the tolerance to which the nonlinear residual was
converged as an additional means of comparison for correctness.

Given the difficulty of these benchmark problems, the linear solve at
each Newton step was preconditioned with an algebraic multigrid
method\footnote{The ML algebraic multigrid library in the Trilinos
  scientific computing libraries \cite{heroux_overview_2005} was used
  for all problems.}, a common technique for preconditioning these
equations \cite{ghia_high-re_1982,evans_enhanced_2007} with the
preconditioner reconstructed at every Newton step for the new
Jacobian. Without multigrid preconditioning, the residual in MCSA was
observed to stagnate. Typically this happens because the Richardson
iteration and Neumann series on which the Monte Carlo method is based
will quickly damp the higher order error modes on the scale of the
computational grid used to solve the problem but will reduce lower
order error modes at a considerably lower rate. In terms of Monte
Carlo, this stagnation manifests itself in histories that simply do no
terminate and therefore no solution is acquired. For many of the
problems presented here, it is likely that a more physics-based
preconditioning strategy such as those presented in
\cite{evans_development_2006,evans_enhanced_2007} will yield better
results for more difficult problems. However, the explicit
preconditioning strategy used here currently limits us to algebraic
preconditioning.

For every benchmark calculation, both FANM and Newton-Krylov were
preconditioned with the same multigrid parameters such that the
conditioning of the system would not be a factor in the verification
of the method or the performance comparison in the following
section. With FANM, this preconditioning was required to reduce the
spectral radius of the Jacobian operators generated at each Newton
step. This preconditioning was also required to obtain good
convergence properties for the Newton-Krylov solutions although it was
not necessary for their convergence. In addition, the same nonlinear
solver parameters including globalization and forcing term strategies
were used with both methods. As with the radiation transport
calculations in the previous chapter, algebraic preconditioning with
the multigrid method results in dense linear systems. To mitigate
this, the reduced domain approximation was again applied to the Monte
Carlo solution within the MCSA solve occurring at each nonlinear
iteration. The values used for the approximation will be given for
each benchmark problem.

Results in this and the following section were produced by the Drekar
multiphysics code base being developed at Sandia National Laboratories
for fluid flow and other multiple physics simulations
\cite{pawlowski_drekar_2012}. Drekar is a massively parallel physics
framework that utilizes Newton-Krylov methods to generate a fully
consistent solution scheme for the Navier-Stokes equations among other
systems with Jacobian generation through automatic differentiation. To
achieve these results, a Monte Carlo solver generated for this work
was injected into the nonlinear tool chain within Drekar to form an
implementation of the FANM method. For the Newton-Krylov solutions,
the same production implementation of GMRES from the Trilinos
scientific computing library Aztec \cite{heroux_overview_2005} used
for the verification of MCSA for the $SP_N$ solutions in the previous
chapter was also used here as a means of comparison and FANM
benchmarking.

@@

For the verification calculations, at each Rayleigh number the global
minimum, maximum, and average values were recorded for fluid velocity
in the x and y directions, fluid pressure, and fluid
temperature. Tables~\ref{tab:convection_ra1e3_results},
\ref{tab:convection_ra1e4_results},
\ref{tab:convection_ra1e5_results}, and
\ref{tab:convection_ra1e6_results} give the results of these
computations. It should be noted that the negative pressures computed
are due to a relative pressure being reported. For each of the data
points presented, both Newton-Krylov solutions and FANM solutions
computed identical numerical results for all of these global
values. No variation in solution was noted for each of the
computations larger than the tolerance of the converged nonlinear
residual norms given in
Table~\ref{tab:convection_residual_norm_comparison}. Variations in the
convergence behavior and differences in performance for these
solutions will be analyzed in the following section.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & -1.288835730E-05 & -1.288835730E-05 \\
      Minimum $U$ & -3.677243960E+00 & -3.677243960E+00 \\
      Maximum $U$ & 3.688908550E+00 & 3.688908550E+00 \\
      \hline
      Average $V$ & 3.403670530E-02 & 3.403670530E-02 \\
      Minimum $V$ & -3.721178580E+00 & -3.721178580E+00 \\
      Maximum $V$ & 3.791483420E+00 & 3.791483420E+00 \\
      \hline
      Average $P$ & 2.345554770E+02 & 2.345554770E+02 \\
      Minimum $P$ & -5.509779180E+00 & -5.509779180E+00 \\
      Maximum $P$ & 4.994885560E+02 & 4.994885560E+02 \\
      \hline
      Average $T$ & 3.595148700E-02 & 3.595148700E-02 \\
      Minimum $T$ & 2.092448340E-04 & 2.092448340E-04 \\
      Maximum $T$ & 7.178874510E-02 & 7.178874510E-02 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity FANM verification for a
      Rayleigh number of \sn{1}{3}.} \textit{The values reported are
      global minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:convection_ra1e3_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & -7.489201240E-04 & -7.489201240E-04 \\
      Minimum $U$ & -1.607318000E+01 & -1.607318000E+01 \\
      Maximum $U$ & 1.669098270E+01 & 1.669098270E+01 \\
      \hline
      Average $V$ & 3.377624420E-01 & 3.377624420E-01 \\
      Minimum $V$ & -1.932174800E+01 & -1.932174800E+01 \\
      Maximum $V$ & 2.047326820E+01 & 2.047326820E+01 \\
      \hline
      Average $P$ & 1.861894950E+03 & 1.861894950E+03 \\
      Minimum $P$ & -3.353416240E+01 & -3.353416240E+01 \\
      Maximum $P$ & 4.440642810E+03 & 4.440642810E+03 \\
      \hline
      Average $T$ & 3.510568440E-01 & 3.510568440E-01 \\
      Minimum $T$ & 2.092448340E-04 & 2.092448340E-04 \\
      Maximum $T$ & 7.178874510E-02 & 7.178874510E-02 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity FANM verification for a
      Rayleigh number of \sn{1}{4}.} \textit{The values reported are
      global minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:convection_ra1e4_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & -2.043639370E-02 & -2.043639370E-02 \\
      Minimum $U$ & -3.814922470E+01 & -3.814922470E+01 \\
      Maximum $U$ & -3.814922470E+01 & -3.814922470E+01 \\
      \hline
      Average $V$ & 2.774864300E+00 & 2.774864300E+00 \\
      Minimum $V$ & -6.038682990E+01 & -6.038682990E+01 \\
      Maximum $V$ & 7.873873560E+01 & 7.873873560E+01 \\
      \hline
      Average $P$ & 1.324208380E+04 & 1.324208380E+04 \\
      Minimum $P$ & -1.390235490E+02 & -1.390235490E+02 \\
      Maximum $P$ & 3.559215530E+04 & 3.559215530E+04 \\
      \hline
      Average $T$ & 3.027827770E+00 & 3.027827770E+00 \\
      Minimum $T$ & 1.626405740E-02 & 1.626405740E-02 \\
      Maximum $T$ & 7.162571130E+00 & 7.162571130E+00 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity FANM verification for a
      Rayleigh number of \sn{1}{5}.} \textit{The values reported are
      global minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:convection_ra1e5_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & -1.802569840E-01 & -1.802569840E-01 \\
      Minimum $U$ & -6.292084130E+01 & -6.292084130E+01 \\
      Maximum $U$ & 2.951113240E+02 & 2.951113240E+02 \\
      \hline
      Average $V$ & 1.336398550E+01 & 1.336398550E+01 \\
      Minimum $V$ & -1.345663290E+02 & -1.345663290E+02 \\
      Maximum $V$ & 3.227873890E+02 & 3.227873890E+02 \\
      \hline
      Average $P$ & 6.354313840E+04 & 6.354313840E+04 \\
      Minimum $P$ & -1.356309630E+02 & -1.356309630E+02 \\
      Maximum $P$ & 2.316308730E+05 & 2.316308730E+05 \\
      \hline
      Average $T$ & 1.780727610E+01 & 1.780727610E+01 \\
      Minimum $T$ & 8.265788890E-02 & 8.265788890E-02 \\
      Maximum $T$ & 7.098873170E+01 & 7.098873170E+01 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity FANM verification for a
      Rayleigh number of \sn{1}{6}.} \textit{The values reported are
      global minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:convection_ra1e6_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{ccc}\hline\hline
      \multicolumn{1}{c}{Rayleigh Number}& 
      \multicolumn{1}{c}{Newton Krylov $||\ve{F(\ve{u})}||$}&
      \multicolumn{1}{c}{FANM $||\ve{F(\ve{u})}||$}\\
      \hline
      \sn{1}{3} & \sn{4.542}{-14} & \sn{1.208}{-14} \\
      \sn{1}{4} & \sn{1.045}{-12} & \sn{7.012}{-13} \\
      \sn{1}{5} & \sn{1.784}{-12} & \sn{1.059}{-12} \\
      \sn{1}{6} & \sn{3.404}{-12} & \sn{3.479}{-12} \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Thermal convection cavity nonlinear residual norm
      achieved convergence.} \textit{The results presented here were
      obtained from the benchmark verification calculations.}}
  \label{tab:convection_residual_norm_comparison}
\end{table}

@@

For the verification calculations, at each Reynolds number the global
minimum, maximum, and average values were recorded for the fluid
velocity in the x and y directions and the fluid
pressure. Tables~\ref{tab:driven_re100_results},
\ref{tab:driven_re300_results}, \ref{tab:driven_re500_results}, and
\ref{tab:driven_re700_results} give the results of these
computations. As with the thermal convection problem, the negative
pressures computed are due to a relative pressure being reported. For
each of the data points given, both Newton-Krylov solutions and FANM
solutions computed identical numerical results for all of these global
values. The norm of the converged nonlinear residual for the
calculations is given in
Table~\ref{tab:driven_residual_norm_comparison}. All computed global
values were within these tolerances.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 1.165632320E-01 & 1.165632320E-01 \\
      Minimum $U$ & -2.423642920E+01 & -2.423642920E+01 \\
      Maximum $U$ & 9.742670880E+01 & 9.742670880E+01 \\
      \hline
      Average $V$ & 2.038480420E-02 & 2.038480420E-02 \\
      Minimum $V$ & -5.266315430E+01 & -5.266315430E+01 \\
      Maximum $V$ & 2.720044160E+01 & 2.720044160E+01 \\
      \hline
      Average $P$ & -1.656193900E+02 & -1.656193900E+02 \\
      Minimum $P$ & -9.635479370E+03 & -9.635479370E+03 \\
      Maximum $P$ & 1.854517020E+04 & 1.854517020E+04 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity FANM verification for a Reynolds
      number of 100.} \textit{The values reported are global minima,
      maxima, and averages for each variable in the problem. The
      velocity is given in components in the $x$ and $y$ directions.}}
  \label{tab:driven_re100_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 3.382078040E-01 & 3.382078040E-01 \\
      Minimum $U$ & -9.467140360E+01 & -9.467140360E+01 \\
      Maximum $U$ & 2.895885380E+02 & 2.895885380E+02 \\
      \hline
      Average $V$ & 8.780242440E-02 & 8.780242440E-02 \\
      Minimum $V$ & -1.830156060E+02 & -1.830156060E+02 \\
      Maximum $V$ & 8.827373340E+01 & 8.827373340E+01 \\
      \hline
      Average $P$ & -2.830380160E+03 & -2.830380160E+03 \\
      Minimum $P$ & -2.352177540E+04 & -2.352177540E+04 \\
      Maximum $P$ & 9.254997120E+04 & 9.254997120E+04 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity FANM verification for a Reynolds
      number of 300.} \textit{The values reported are global minima,
      maxima, and averages for each variable in the problem. The
      velocity is given in components in the $x$ and $y$ directions.}}
  \label{tab:driven_re300_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 5.098921320E-01 & 5.098921320E-01 \\
      Minimum $U$ & -1.731861690E+02 & -1.731861690E+02 \\
      Maximum $U$ & 4.789146040E+02 & 4.789146040E+02 \\
      \hline
      Average $V$ & 1.365493140E-01 & 1.365493140E-01 \\
      Minimum $V$ & -3.170930360E+02 & -3.170930360E+02 \\
      Maximum $V$ & 1.655401440E+02 & 1.655401440E+02 \\
      \hline
      Average $P$ & -9.378317550E+03 & -9.378317550E+03 \\
      Minimum $P$ & -4.155215280E+04 & -4.155215280E+04 \\
      Maximum $P$ & 2.058742010E+05 & 2.058742010E+05 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity FANM verification for a Reynolds
      number of 500.} \textit{The values reported are global minima,
      maxima, and averages for each variable in the problem. The
      velocity is given in components in the $x$ and $y$ directions.}}
  \label{tab:driven_re500_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 6.479541170E-01 & 6.479541170E-01 \\
      Minimum $U$ & -2.550685400E+02 & -2.550685400E+02 \\
      Maximum $U$ & 6.663054900E+02 & 6.663054900E+02 \\
      \hline
      Average $V$ & 1.833690840E-01 & 1.833690840E-01 \\
      Minimum $V$ & -4.491052980E+02 & -4.491052980E+02 \\
      Maximum $V$ & 2.474376340E+02 & 2.474376340E+02 \\
      \hline
      Average $P$ & 1.833690840E-01 & 1.833690840E-01 \\
      Minimum $P$ & -6.241704500E+04 & -6.241704500E+04 \\
      Maximum $P$ & 3.518174810E+05 & 3.518174810E+05 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity FANM verification for a Reynolds
      number of 700.} \textit{The values reported are global minima,
      maxima, and averages for each variable in the problem. The
      velocity is given in components in the $x$ and $y$ directions.}}
  \label{tab:driven_re700_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{l}{Reynolds Number}& 
      \multicolumn{1}{c}{Newton Krylov $||\ve{F(\ve{u})}||$}&
      \multicolumn{1}{c}{FANM $||\ve{F(\ve{u})}||$}\\
      \hline
      100 & \sn{5.453}{-14} & \sn{6.577}{-14} \\
      300 & \sn{2.537}{-13} & \sn{2.779}{-13} \\
      500 & \sn{6.367}{-13} & \sn{6.252}{-13} \\
      700 & \sn{9.159}{-13} & \sn{1.282}{-12} \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Lid driven cavity nonlinear residual norm achieved
      convergence.} \textit{The results presented here were obtained
      from the benchmark verification calculations.}}
  \label{tab:driven_residual_norm_comparison}
\end{table}

@@

For the verification calculations, at each Reynolds number the global
minimum, maximum, and average values were recorded for the fluid
velocity in the x and y directions and the fluid
pressure. Tables~\ref{tab:step_re200_results},
\ref{tab:step_re300_results}, \ref{tab:step_re400_results}, and
\ref{tab:step_re500_results} give the results of these computations.
The norm of the converged nonlinear residual for the calculations is
given in Table~\ref{tab:step_residual_norm_comparison}. The
Newton-Krylov solutions and FANM solutions computed identical
numerical results for all of these global values within converged
tolerances.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 8.776187720E+01 & 8.776187720E+01 \\
      Minimum $U$ & -2.796828970E+01 & -2.796828970E+01 \\
      Maximum $U$ & 2.639993990E+02 & 2.639993990E+02 \\
      \hline
      Average $V$ & -8.870766180E-01 & -8.870766180E-01 \\
      Minimum $V$ & -2.636357380E+01 & -2.636357380E+01 \\
      Maximum $V$ & 1.330345700E+01 & 1.330345700E+01 \\
      \hline
      Average $P$ & 1.226960220E+04 & 1.226960220E+04 \\ 
      Minimum $P$ & 2.953808400E+01 & 2.953808400E+01 \\ 
      Maximum $P$ & 2.202121190E+04 & 2.202121190E+04 \\ 
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step FANM verification for a
      Reynolds number of 200.} \textit{The values reported are global
      minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:step_re200_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 1.316701890E+02 & 1.316701890E+02 \\
      Minimum $U$ & -4.357431070E+01 & -4.357431070E+01 \\
      Maximum $U$ & 3.958553050E+02 & 3.958553050E+02 \\
      \hline
      Average $V$ & -1.330267460E+00 & -1.330267460E+00 \\
      Minimum $V$ & -3.467114630E+01 & -3.467114630E+01 \\
      Maximum $V$ & 1.694243270E+01 & 1.694243270E+01 \\
      \hline
      Average $P$ & 1.681552980E+04 & 1.681552980E+04 \\
      Minimum $P$ & 4.129311640E+01 & 4.129311640E+01 \\
      Maximum $P$ & 2.940820620E+04 & 2.940820620E+04 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step FANM verification for a
      Reynolds number of 300.} \textit{The values reported are global
      minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:step_re300_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 1.759720450E+02 & 1.759720450E+02 \\
      Minimum $U$ & -5.941731260E+01 & -5.941731260E+01 \\
      Maximum $U$ & 5.287146290E+02 & 5.287146290E+02 \\
      \hline
      Average $V$ & -1.776706740E+00 & -1.776706740E+00 \\
      Minimum $V$ & -4.497362670E+01 & -4.497362670E+01 \\
      Maximum $V$ & 1.973372840E+01 & 1.973372840E+01 \\
      \hline
      Average $P$ & 1.954352940E+04 & 1.954352940E+04 \\
      Minimum $P$ & 5.294095080E+01 & 5.294095080E+01 \\
      Maximum $P$ & 3.457969480E+04 & 3.457969480E+04 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step FANM verification for a
      Reynolds number of 400.} \textit{The values reported are global
      minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:step_re400_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lll}\hline\hline
      \multicolumn{1}{l}{Parameter}& 
      \multicolumn{1}{l}{Newton-Krylov Result}&
      \multicolumn{1}{l}{FANM Result}\\
      \hline
      Average $U$ & 2.199858290E+02 & 2.199858290E+02 \\
      Minimum $U$ & -7.549221730E+01 & -7.549221730E+01 \\
      Maximum $U$ & 6.605863330E+02 & 6.605863330E+02 \\
      \hline
      Average $V$ & -2.220108250E+00 & -2.220108250E+00 \\
      Minimum $V$ & -5.888473080E+01 & -5.888473080E+01 \\
      Maximum $V$ & 2.201938360E+01 & 2.201938360E+01 \\
      \hline
      Average $P$ & 1.983947940E+04 & 1.983947940E+04 \\
      Minimum $P$ & -9.707913370E+03 & -9.707913370E+03 \\
      Maximum $P$ & 3.771193860E+04 & 3.771193860E+04 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step FANM verification for a
      Reynolds number of 500.} \textit{The values reported are global
      minima, maxima, and averages for each variable in the
      problem. The velocity is given in components in the $x$ and $y$
      directions.}}
  \label{tab:step_re500_results}
\end{table}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{ccc}\hline\hline
      \multicolumn{1}{c}{Reynolds Number}& 
      \multicolumn{1}{c}{Newton Krylov $||\ve{F(\ve{u})}||$}&
      \multicolumn{1}{c}{FANM $||\ve{F(\ve{u})}||$}\\
      \hline
      200 & 1.939E-11 & 1.975E-11 \\
      300 & 2.786E-11 & 2.751E-11 \\
      400 & 3.360E-11 & 3.597E-11 \\
      500 & 5.824E-11 & 4.062E-11 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Backward facing step nonlinear residual norm
      achieved convergence.} \textit{The results presented here were
      obtained from the benchmark verification calculations.}}
  \label{tab:step_residual_norm_comparison}
\end{table}

>>>>

%%---------------------------------------------------------------------------%%
\section{The FANM Method\ }
\label{sec:fanm}
In physics codes based on nonlinear equations sets, Newton-Krylov
methods are the primary means of generating a fully consistent
solution scheme
\cite{evans_development_2006,evans_enhanced_2007,gaston_parallel_2009,godoy_parallel_2012}. It
is common that for large scale simulations these problems are memory
limited due to the subspaces generated by robust Krylov methods which
may often build hundreds of vectors during a Newton step. Often, a
matrix-free approach is chosen to relax memory requirements over
directly generating the Jacobian matrix and facilitate the
implementation. However, as we observed in previous sections, these
matrix-free methods suffer from poorly scaled problems and the first
order error introduced by the Jacobian approximation. In addition, it
was observed that the savings induced by the matrix-free approach is
eventually amortized over a number of nonlinear iterations where it
becomes more efficient computationally to instead form the Jacobian.

@@

For each of the benchmark problems, we will present the results of
computations using FANM and compare them to results using a
Newton-Krylov method in order to verify the correctness of FANM and
its applicability to fluid problems. Given the difficulty of these
benchmark problems, the linear solve at each Newton step was
preconditioned with an algebraic multigrid method\footnote{The ML
  algebraic multigrid library in the Trilinos scientific computing
  libraries \cite{heroux_overview_2005} was used for all problems.}, a
common technique for preconditioning these equations
\cite{ghia_high-re_1982,evans_enhanced_2007} with the preconditioner
reconstructed at every Newton step for the new Jacobian. Without
multigrid preconditioning, the residual in MCSA was observed to
stagnate. Typically this happens because the Richardson iteration and
Neumann series on which the Monte Carlo method is based will quickly
damp the higher order error modes on the scale of the computational
grid used to solve the problem but will reduce lower order error modes
at a considerably lower rate. In terms of Monte Carlo, this stagnation
manifests itself in histories that simply do no terminate and
therefore no solution is acquired. For many of the problems presented
here, it is likely that a more physics-based preconditioning strategy
such as those presented in
\cite{evans_development_2006,evans_enhanced_2007} will yield better
results for more difficult problems. However, the explicit
preconditioning strategy used here currently limits us to algebraic
preconditioning.

For every benchmark calculation, both FANM and Newton-Krylov were
preconditioned with the same multigrid parameters such that the
conditioning of the system would not be a factor in the verification
of the method or the performance comparison in the following
section. With FANM, this preconditioning was required to reduce the
spectral radius of the Jacobian operators generated at each Newton
step. This preconditioning was also required to obtain good
convergence properties for the Newton-Krylov solutions although it was
not necessary for their convergence. In addition, the same nonlinear
solver parameters including globalization and forcing term strategies
were used with both methods with a nonlinear convergence criteria of a
\sn{1}{-12} norm for the nonlinear residual. As with the radiation
transport calculations in the previous chapter, algebraic
preconditioning with the multigrid method results in dense linear
systems. To mitigate this, the reduced domain approximation was again
applied to the Monte Carlo solution within the MCSA solve occurring at
each nonlinear iteration. The values used for the approximation will
be given for each benchmark problem.

Results in this and the following section were produced by the Drekar
multiphysics code base being developed at Sandia National Laboratories
for fluid flow and other multiple physics simulations
\cite{pawlowski_drekar_2012}. Drekar is a massively parallel physics
framework that utilizes Newton-Krylov methods to generate a fully
consistent solution scheme for the Navier-Stokes equations among other
systems with Jacobian generation through automatic differentiation. To
achieve these results, a Monte Carlo solver generated for this work
was injected into the nonlinear tool chain within Drekar to form an
implementation of the FANM method. For the Newton-Krylov solutions,
the same implementation of GMRES from the Trilinos scientific
computing library Aztec \cite{heroux_overview_2005} used for the
verification of MCSA for the $SP_N$ solutions in the previous chapter
was also used here as a means of comparison and FANM benchmarking.

@@

For the verification calculations, at each Rayleigh number the fluid
velocity in the x and y directions, fluid pressure, and fluid
temperature values calculated Newton-Krylov were compared against the
Newton-Krylov calculations. No variation in solution was noted for
each of the computations larger than the floating point tolerance
reported by Drekar with an absolute difference value of zero observed
at all spatial locations for each quantity.

@@

For the verification calculations, at each Reynolds number the fluid
velocity in the x and y directions and fluid pressure values
calculated Newton-Krylov were compared against the Newton-Krylov
calculations. No variation in solution was noted for each of the
computations larger than the floating point tolerance reported by
Drekar with an absolute difference value of zero observed at all
spatial locations for each quantity.

@@

For the verification calculations, at each Reynolds number the fluid
velocity in the x and y directions and fluid pressure values
calculated Newton-Krylov were compared against the Newton-Krylov
calculations. No variation in solution was noted for each of the
computations larger than the floating point tolerance reported by
Drekar with an absolute difference value of zero observed at all
spatial locations for each quantity.
